Bayesian determination of the effect of a deep eutectic solvent on the structure of lipid monolayers

### 2018-11-06 reports/paper.tex
more ESI referencing

%
\begin{figure}
	\centering
  \includegraphics[width=0.45\textwidth]{figures/dlpc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dmpc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dppc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dmpg_ref_sld}
	\caption{\small The XRR profiles (left) and SLD profiles (right) for each
  of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG, at the four
  measured surface pressures; see legend above each plot. The different
  surface pressure XRR profiles have been offset in the $y$-axis by an order
  of magnitude and SLD profiles offset in the $y$-axis by
  \SI{5e-6}{\per\square\angstrom}, for clarity \cite{SI2018}.}
	\label{fig:lipids}
\end{figure}
%
Our model has more variables than we can uniquely fit with limited data.
In reflectometry the usual approach to this problem is to collect more
equivalent dataset with varying deuteration (NR). For reasons of cost and
practicality this was not possible in this work.
The other approach is to constrain the model so that the number of fitted
variables are reduced, and is the approach taken here.
To do this we add the constraint that a single tail volume can be used for
all surface pressures. To justify this we need to be sure that the lipids
remain in the same phase. On water this is can be demonstrated with a
Langmuir isotherm.
However, while we have confidence that the individual surface pressures
measured were reliable, we were unable to collect consistent Langmuir
isotherm measurements, due to the high viscosity of the DES.
This inconsistency means that we cannot be confident in the phase of the
lipids based on the surface pressure alone.
Instead we have used grazing incidence X-ray diffraction to confirm the
phases of DMPC and DPPC at \SI{30}{\milli\newton\per\meter}.
DPPC was found to be in the LC phase and DMPC in the LE phase at room
temperature for the surface pressures measured (see Section \ref{sec:gixd}).
We assume that DMPG and DLPC are also in the LE phase since there is no
reason to believe that the phase behaviour in these systems differs
significantly from DMPC at the same temperature.

@@

%
\begin{table}
	\caption{\label{tab:liptab} The best-fit values, and associated 95 \%
  confidence intervals for the varying parameters in the XRR models, at the
  \SI{30}{\milli\newton\per\meter}. The values of $d_t$ were found from the
  appropriate values of $\theta_t$ using Eqn. \ref{equ:tl} and the values for
  $\phi_h$ were obtained from the appropriate use of Eqn. \ref{equ:phih}.}
  \begin{ruledtabular}
	\begin{tabular}{ccccc}
    Lipid & DLPC & DMPC & DPPC & DMPG \\
    \hline
    $\theta_t$/\si{\degree} & \input{../output/dlpc/angle30.txt} &
    \input{../output/dmpc/angle30.txt} & \input{../output/dppc/angle30.txt} &
    \input{../output/dmpg/angle30.txt} \\
    $\sigma$/\si{\angstrom} & \input{../output/dlpc/rough30.txt} &
    \input{../output/dmpc/rough30.txt} & \input{../output/dppc/rough30.txt} &
    \input{../output/dmpg/rough30.txt} \\
    \hline
    $V_t$/\si{\cubic\angstrom} & \input{../output/dlpc/vt.txt} &
    \input{../output/dmpc/vt.txt} & \input{../output/dppc/vt.txt} &
    \input{../output/dmpg/vt.txt} \\
    $V_h$/\si{\cubic\angstrom} & \input{../output/dlpc/vh.txt} &
    \input{../output/dmpc/vh.txt} & \input{../output/dppc/vh.txt} &
    \input{../output/dmpg/vh.txt} \\
    $d_h$/\si{\angstrom} & \input{../output/dlpc/head.txt} &
    \input{../output/dmpc/head.txt} & \input{../output/dppc/head.txt} &
    \input{../output/dmpg/head.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dlpc/solh30.txt} &
    \input{../output/dmpc/solh30.txt} & \input{../output/dppc/solh30.txt} &
    \input{../output/dmpg/solh30.txt} \\
    $d_t$/\si{\angstrom} & \input{../output/dlpc/tail30.txt} &
    \input{../output/dmpc/tail30.txt} & \input{../output/dppc/tail30.txt} &
    \input{../output/dmpg/tail30.txt} \\
	\end{tabular}
  \end{ruledtabular}
\end{table}
%

@@

Our fits suggest lower lipid tail volumes than previous measurements using
other techniques (compare Tables \ref{tab:water} and \ref{tab:liptab}).
It is unlikely that this is a result of the DES subphase since the tails do
not directly interact with the solvent.
However this may be related to the compaction of the monolayer at elevated
surface pressures.
The optimal value of the tail volume for DPPC in the LC
phase\cite{Campbell2018} was found to be \SI{772}{\cubic\angstrom} at
\SI{35}{\milli\newton\per\meter}, which agrees well with the value of
\input{../output/dppc/vt.txt}\si{\cubic\angstrom} found in this work.
Similarly the tail volume for DMPS in the LE phase\cite{Campbell2018},
\SI{714}{\cubic\angstrom} at \SI{10}{\milli\newton\per\meter}, broadly
agrees with the values for
DMPC (\input{../output/dmpc/vt.txt}\si{\cubic\angstrom}) and DMPG
(\input{../output/dmpg/vt.txt}\si{\cubic\angstrom}) in this work.
The literature value\cite{Pan2012} quoted for DMPG (at a lower temperature) is
also similar to our result (but slightly smaller).
We find that the reduction is \SIrange{8}{12}{\percent} for DPPC, DMPC and
DLPC when compared with literature sources at \SIrange{24}{30}{\celsius}, in
good agreement with the maximum compression percentage of \SI{15}{\percent}
noted by Small \emph{et al.}\cite{Small1984}.
In general our results are at least self-consistent while comparisons to the
literature are favourable.

Figure \ref{fig:lipresults} shows the PDFs for the head volume for each of
the four lipids.
The three lipids with the PC head are consistent with values of around
\SI{330}{\cubic\angstrom}, regardless of hydrocarbon tail.
This agrees well with the values found for the same head in water
(Table \ref{tab:water}).
Interestingly, the volume for the PG head is similar to that for the PC head
with a value of \input{../output/dmpg/vh.txt}\si{\cubic\angstrom}, which is
significantly larger than values quoted in the literature for
DMPG\cite{Pan2012} (\SI{291}{\cubic\angstrom}) or POPG\cite{Kucerka2012}
(\SI{289}{\cubic\angstrom}).
This suggests that there may be some effect arising from the solvation in DES.

@@

The volumes determined for the head and tails, and the thickness of the head
layers, for DMPC and DPPC were used to constrain the fitting of corresponding
NR data (Figure \ref{fig:neutron}). This leaves only two variables,
$\theta_t$ and $\sigma$ to fit this data.
The agreement is good and a confirmation that the volumes derived from XRR
are consistent with the NR data.
Furthermore, the observed trends with surface pressure are also consistent
which adds some confidence in our conclusions.
%
\begin{figure}
	\centering
  \includegraphics[width=0.45\textwidth]{figures/dmpc_20n_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dppc_20n_ref_sld}
	\caption{\small The NR and SLD profiles at a surface pressure of
  20 mNm$^{-1}$ for two contrasts (see legend above each plot); (a) DMPC,
  (b) DPPC. The NR profiles have been offset in the $y$-axis by an order of
  magnitude and SLD profiles offset in the $y$-axis by
  \SI{5e-6}{\per\square\angstrom}, for clarity \cite{SI2018}.}
	\label{fig:neutron}
\end{figure}
%
%
\begin{table}
  \caption{\label{tab:neutron} The best-fit values, and associated 95 \%
  confidence intervals for the varying parameters in the co-refined NR models.
  The values of $d_t$ were found from the appropriate values of $\theta_t$
  using Eqn. \ref{equ:tl}, and the values of $\phi_h$ were found using Eqn.
  \ref{equ:phih}.}
  \begin{ruledtabular}
	\begin{tabular}{ccccc}
    Lipid & \multicolumn{2}{c}{d$_{54}$-DMPC} & \multicolumn{2}{c}{d$_{62}$-DPPC} \\
    SP/mNm$^{-1}$ & 20 & 25 & 15 & 20 \\
    \hline
    $\theta_t$/$^\circ$ & \input{../output/dmpc/angle20_neutron.txt} &
    \input{../output/dmpc/angle25_neutron.txt} &
    \input{../output/dppc/angle15_neutron.txt} &
    \input{../output/dppc/angle20_neutron.txt} \\
    $\sigma_{t,h,s}$/\AA & \input{../output/dmpc/rough20_neutron.txt} &
    \input{../output/dmpc/rough25_neutron.txt} &
    \input{../output/dppc/rough15_neutron.txt} &
    \input{../output/dppc/rough20_neutron.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dmpc/solh20_neutron.txt} &
    \input{../output/dmpc/solh25_neutron.txt} &
    \input{../output/dppc/solh15_neutron.txt} &
    \input{../output/dppc/solh20_neutron.txt} \\
    $d_t$/\AA & \input{../output/dmpc/tail20_neutron.txt} &
    \input{../output/dmpc/tail25_neutron.txt} &
    \input{../output/dppc/tail15_neutron.txt} &
    \input{../output/dppc/tail20_neutron.txt} \\
	\end{tabular}
  \end{ruledtabular}
\end{table}
%

>>>>

%
\begin{figure}
	\centering
  \includegraphics[width=0.45\textwidth]{figures/dlpc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dmpc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dppc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dmpg_ref_sld}
	\caption{\small The XRR profiles (left) and SLD profiles (right) for each
  of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG, at the four
  measured surface pressures; see legend above each plot. The different
  surface pressure XRR profiles have been offset in the $y$-axis by an order
  of magnitude and SLD profiles offset in the $y$-axis by
  \SI{5e-6}{\per\square\angstrom}, for clarity \cite{SI2018}.}
	\label{fig:lipids}
\end{figure}
%
Our model has more variables than we can uniquely fit with limited data.
In reflectometry the usual approach to this problem is to collect more
equivalent dataset with varying deuteration (NR). For reasons of cost and
practicality this was not possible in this work.
The other approach is to constrain the model so that the number of fitted
variables are reduced, and is the approach taken here.
To do this we add the constraint that a single tail volume can be used for
all surface pressures. To justify this we need to be sure that the lipids
remain in the same phase. On water this is can be demonstrated with a
Langmuir isotherm.
However, while we have confidence that the individual surface pressures
measured were reliable, we were unable to collect consistent Langmuir
isotherm measurements, due to the high viscosity of the DES.
This inconsistency means that we cannot be confident in the phase of the
lipids based on the surface pressure alone.
Instead we have used grazing incidence X-ray diffraction to confirm the
phases of DMPC and DPPC at \SI{30}{\milli\newton\per\meter}.
DPPC was found to be in the LC phase and DMPC in the LE phase at room
temperature for the surface pressures measured (see Section \ref{sec:gixd}) \cite{SI2018}.
We assume that DMPG and DLPC are also in the LE phase since there is no
reason to believe that the phase behaviour in these systems differs
significantly from DMPC at the same temperature.

@@

%
\begin{table}
	\caption{\label{tab:liptab} The best-fit values, and associated 95 \%
  confidence intervals for the varying parameters in the XRR models, at the
  \SI{30}{\milli\newton\per\meter}. The values of $d_t$ were found from the
  appropriate values of $\theta_t$ using Eqn. \ref{equ:tl} and the values for
  $\phi_h$ were obtained from the appropriate use of Eqn. \ref{equ:phih} \cite{SI2018}.}
  \begin{ruledtabular}
	\begin{tabular}{ccccc}
    Lipid & DLPC & DMPC & DPPC & DMPG \\
    \hline
    $\theta_t$/\si{\degree} & \input{../output/dlpc/angle30.txt} &
    \input{../output/dmpc/angle30.txt} & \input{../output/dppc/angle30.txt} &
    \input{../output/dmpg/angle30.txt} \\
    $\sigma$/\si{\angstrom} & \input{../output/dlpc/rough30.txt} &
    \input{../output/dmpc/rough30.txt} & \input{../output/dppc/rough30.txt} &
    \input{../output/dmpg/rough30.txt} \\
    \hline
    $V_t$/\si{\cubic\angstrom} & \input{../output/dlpc/vt.txt} &
    \input{../output/dmpc/vt.txt} & \input{../output/dppc/vt.txt} &
    \input{../output/dmpg/vt.txt} \\
    $V_h$/\si{\cubic\angstrom} & \input{../output/dlpc/vh.txt} &
    \input{../output/dmpc/vh.txt} & \input{../output/dppc/vh.txt} &
    \input{../output/dmpg/vh.txt} \\
    $d_h$/\si{\angstrom} & \input{../output/dlpc/head.txt} &
    \input{../output/dmpc/head.txt} & \input{../output/dppc/head.txt} &
    \input{../output/dmpg/head.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dlpc/solh30.txt} &
    \input{../output/dmpc/solh30.txt} & \input{../output/dppc/solh30.txt} &
    \input{../output/dmpg/solh30.txt} \\
    $d_t$/\si{\angstrom} & \input{../output/dlpc/tail30.txt} &
    \input{../output/dmpc/tail30.txt} & \input{../output/dppc/tail30.txt} &
    \input{../output/dmpg/tail30.txt} \\
	\end{tabular}
  \end{ruledtabular}
\end{table}
%

@@

Our fits suggest lower lipid tail volumes than previous measurements using
other techniques (compare Tables \ref{tab:water} and \ref{tab:liptab}) \cite{SI2018}.
It is unlikely that this is a result of the DES subphase since the tails do
not directly interact with the solvent.
However this may be related to the compaction of the monolayer at elevated
surface pressures.
The optimal value of the tail volume for DPPC in the LC
phase\cite{Campbell2018} was found to be \SI{772}{\cubic\angstrom} at
\SI{35}{\milli\newton\per\meter}, which agrees well with the value of
\input{../output/dppc/vt.txt}\si{\cubic\angstrom} found in this work.
Similarly the tail volume for DMPS in the LE phase\cite{Campbell2018},
\SI{714}{\cubic\angstrom} at \SI{10}{\milli\newton\per\meter}, broadly
agrees with the values for
DMPC (\input{../output/dmpc/vt.txt}\si{\cubic\angstrom}) and DMPG
(\input{../output/dmpg/vt.txt}\si{\cubic\angstrom}) in this work.
The literature value\cite{Pan2012} quoted for DMPG (at a lower temperature) is
also similar to our result (but slightly smaller).
We find that the reduction is \SIrange{8}{12}{\percent} for DPPC, DMPC and
DLPC when compared with literature sources at \SIrange{24}{30}{\celsius}, in
good agreement with the maximum compression percentage of \SI{15}{\percent}
noted by Small \emph{et al.}\cite{Small1984}.
In general our results are at least self-consistent while comparisons to the
literature are favourable.

Figure \ref{fig:lipresults} shows the PDFs for the head volume for each of
the four lipids.
The three lipids with the PC head are consistent with values of around
\SI{330}{\cubic\angstrom}, regardless of hydrocarbon tail.
This agrees well with the values found for the same head in water
(Table \ref{tab:water}) \cite{SI2018}.
Interestingly, the volume for the PG head is similar to that for the PC head
with a value of \input{../output/dmpg/vh.txt}\si{\cubic\angstrom}, which is
significantly larger than values quoted in the literature for
DMPG\cite{Pan2012} (\SI{291}{\cubic\angstrom}) or POPG\cite{Kucerka2012}
(\SI{289}{\cubic\angstrom}).
This suggests that there may be some effect arising from the solvation in DES.

@@

The volumes determined for the head and tails, and the thickness of the head
layers, for DMPC and DPPC were used to constrain the fitting of corresponding
NR data (Figure \ref{fig:neutron}). This leaves only two variables,
$\theta_t$ and $\sigma$ to fit this data.
The agreement is good and a confirmation that the volumes derived from XRR
are consistent with the NR data.
Furthermore, the observed trends with surface pressure are also consistent
which adds some confidence in our conclusions.
%
\begin{figure}
	\centering
  \includegraphics[width=0.45\textwidth]{figures/dmpc_20n_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dppc_20n_ref_sld}
	\caption{\small The NR and SLD profiles at a surface pressure of
  20 mNm$^{-1}$ for two contrasts (see legend above each plot); (a) DMPC,
  (b) DPPC. The NR profiles have been offset in the $y$-axis by an order of
  magnitude and SLD profiles offset in the $y$-axis by
  \SI{5e-6}{\per\square\angstrom}, for clarity \cite{SI2018}.}
	\label{fig:neutron}
\end{figure}
%
%
\begin{table}
  \caption{\label{tab:neutron} The best-fit values, and associated 95 \%
  confidence intervals for the varying parameters in the co-refined NR models.
  The values of $d_t$ were found from the appropriate values of $\theta_t$
  using Eqn. \ref{equ:tl}, and the values of $\phi_h$ were found using Eqn.
  \ref{equ:phih} \cite{SI2018}.}
  \begin{ruledtabular}
	\begin{tabular}{ccccc}
    Lipid & \multicolumn{2}{c}{d$_{54}$-DMPC} & \multicolumn{2}{c}{d$_{62}$-DPPC} \\
    SP/mNm$^{-1}$ & 20 & 25 & 15 & 20 \\
    \hline
    $\theta_t$/$^\circ$ & \input{../output/dmpc/angle20_neutron.txt} &
    \input{../output/dmpc/angle25_neutron.txt} &
    \input{../output/dppc/angle15_neutron.txt} &
    \input{../output/dppc/angle20_neutron.txt} \\
    $\sigma_{t,h,s}$/\AA & \input{../output/dmpc/rough20_neutron.txt} &
    \input{../output/dmpc/rough25_neutron.txt} &
    \input{../output/dppc/rough15_neutron.txt} &
    \input{../output/dppc/rough20_neutron.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dmpc/solh20_neutron.txt} &
    \input{../output/dmpc/solh25_neutron.txt} &
    \input{../output/dppc/solh15_neutron.txt} &
    \input{../output/dppc/solh20_neutron.txt} \\
    $d_t$/\AA & \input{../output/dmpc/tail20_neutron.txt} &
    \input{../output/dmpc/tail25_neutron.txt} &
    \input{../output/dppc/tail15_neutron.txt} &
    \input{../output/dppc/tail20_neutron.txt} \\
	\end{tabular}
  \end{ruledtabular}
\end{table}
%
